SciPy 1.0---Fundamental Algorithms for Scientific Computing in Python

### 2018-07-26 scipy-1.0/paper.tex
Update scipy.stats section for review comments.

\textbf{Project scope}
SciPy aims to provide fundamental algorithms for scientific computing. The
breadth of its scope was derived from the Guide to Available Mathematical
Software classification system (GAMS\cite{boisvert1991guide}). In areas
that move relatively slowly, e.g. linear algebra, SciPy aims to provide
complete coverage. In other areas it aims to provide fundamental building
blocks, while interacting well with other packages specialized in that area.
For example for statistics, SciPy provides what one expect to find in a
statistics textbook (probability distributions, hypothesis tests, frequency
statistics, correlation functions, and more), while Statsmodels provides
more advanced statistical estimators and inference methods, scikit-learn
covers machine learning, PyMC3, emcee and PyStan cover Bayesian statistics
and probabilistic modelling.

We use the following criteria to determine whether or not to include new
functionality in SciPy:
\begin{itemize}
    \item The algorithm is of relevance to multiple fields of science.
    \item The implementation is demonstrably important.  E.g., it is classic
    enough to be included in textbooks, or it is based on a peer-reviewed article
    which has a significant number of citations.
\end{itemize}

In terms of software systems and architecture, SciPy's scope matches NumPy's:
algorithms for in-memory computing on single machines, with support for a wide
range of data types and process architectures. Distributed computing and support
for graphics processing units (GPUs) is explicitly out of scope.

@@

\subsection*{API and ABI evolution}
The application programming interface (API) for SciPy consists of
approximately 1500 functions and classes.
Our policy for evolving the API over time is that new functionality can be added,
while removing or changing existing functionality can only be done if the benefits
of that exceeds the (often significant) costs to users, \textit{and} only
after giving clear deprecation warnings to those users for at least one year.
In general, we encourage changes that improve clarity in the API of the
library but strongly discourage breaking backwards compatibility\footnote{
    Backwards compatibility means that new versions of SciPy will keep user
    code written for an older version working.}
given our position near the base of the scientific Python computing stack.

In addition to the Python API, SciPy has C and Cython interfaces in a number
of submodules. Therefore, we have to also consider the application binary
interface (ABI). This ABI has been stable for a long time, and we aim to
evolve it only in a backwards compatible way.

@@

While \texttt{scipy.sparse} is mature, there are many data
formats and the community continues to move forward here so we will aim
to take advantage of their improvements\cite{abbasi2018sparse} 
while preserving backwards compatibility to the extent that it is possible.
\texttt{sparse.linalg} may benefit from wrappers for PROPACK 
for faster sparse SVD computation.

@@

\textbf{Maintainers and contributors}
The SciPy project has approximately 15 active maintainers (people who review
others' contributions, and in general do everything needed to ensure that
the software and the project move forward).  It also has an additional ~100
unique contributors for every 6-monthly release cycle.  Maintainers are
critical to the health of the project\cite{eghbal2016}; they enable contributions
from a much larger group, and their skills and efforts largely determine how
fast the project moves forward.  Anyone with the interest and skills can become
a maintainer; the SciPy Developer Guide provides guidance on how to do that.

@@

From the user side, there were 13,096,468 PyPI downloads
of SciPy during the year 2017\cite{pypinfo}, which is likely a vast
underestimate of total downloads given that we don't have access to the
popular Anaconda package manager download statistics. The recommended 
source for citing SciPy is simply a reference to the website / library
itself\cite{SciPylib}, and has been cited $>3000$ times. Some of the most
prominent usages of or demonstrations of credibility for SciPy include the LIGO / Virgo
scientific collaboration that lead to the observation of gravitational
waves \cite{PhysRevLett.116.061102}, and the fact that SciPy is shipped directly
with MacOS and in the Intel Distribution for Python\cite{intel-python}.

\textbf{Future development}.
Future development plans include a major rewrite of the \texttt{scipy.sparse} module, which is already underway\cite{abbasi2018sparse}. This is due to the current interface only supporting two-dimensional objects. Even sub-arrays are two-dimensional, and dimensions greater than two aren't supported. The new package supports dimensions greater than two, and it bases itself around arrays rather than matrices (which is strictly more powerful). It also features improved inter-operability with NumPy and other community packages.

\textit{This section should include key issues: sparse arrays, ndimage pixel vs point, splines, fftpack vs. np.fft and linalg vs. np.linalg, under-maintained submodules.}

>>>>

\textbf{Project goals and scope}

@@

\subsection*{API and ABI evolution}
In general, we encourage changes that improve clarity in the API of the library,
but strongly discourage breaking backwards compatibility, given our position near
the base of the scientific Python computing stack.

@@

While \texttt{scipy.sparse} is mature, there are many data
formats and the community continues to move forward here so we will aim
to take advantage of their improvements\cite{Sparse_Arrays} 
while preserving backwards compatibility to the extent that it is possible.
\texttt{sparse.linalg} may benefit from wrappers for PROPACK 
for faster sparse SVD computation.

@@

\textbf{Maintainers and contributors}

@@

% need some better usage stats sources?
% SciPy stuff from njs blog: https://docs.google.com/spreadsheets/d/1lOLvSF0up4eZyv2ugZi-TM_GIs3gPCkGNIDKXS1Y3w4/edit#gid=406031940
From the user side, there have been at least 15 million downloads
of the SciPy manylinux wheels between the start of 2016 and May of 2018
(njs blog). The recommended source for citing SciPy is simply a reference
to the website / library itself\cite{SciPylib}, and has been cited $>3000$
times. Some of the most prominent citing articles include the IPython
publication\cite{PER-GRA:2007}, a heavily-cited Nature Biotechnology
article\cite{pmid17287757}, a dark
matter annihilitation study in the field of
cosmology\cite{1475-7516-2012-08-007}, and many of the
other major downstream software libraries in the Python scientific
computing stack.

\textbf{Future development}.
\textit{This section should include key issues: sparse arrays, ndimage pixel vs point, splines, fftpack vs. np.fft and linalg vs. np.linalg, under-maintained submodules.}
